---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load the activity.csv file into R
```{r}
data = read.csv('activity.csv', header = T) #load data
options("scipen" =100, "digits" = 2) # override R's tendency to use scientific notation
library(ggplot2) #load graphics package
```



## What is mean total number of steps taken per day?
Calculate the total steps per day
```{r}
totalsteps <- aggregate(steps ~ date, data, sum)
head(totalsteps)
hist(totalsteps$steps)
```

Calculate the mean and median steps:
```{r}
mean(totalsteps$steps, na.rm=TRUE)
median(totalsteps$steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
timeseries <- aggregate(steps ~ interval, data = data, FUN = mean) #aggregate data
maxsteps <- timeseries[which.max(timeseries$steps),]$interval #identify interval with max steps

plot(timeseries, type="l",xlab="Interval",ylab="Average steps per day",
     main="Average daily activity pattern")
abline(v = maxsteps , col = "red")
```

The maximum number of steps occur during interval **`r maxsteps`**.


## Imputing missing values
1. ***Calculate the number of missing values:***

```{r}
missing = sum(is.na(data$steps))
```

The total number of missing values is **`r missing`**.


2. ***Fill in the missing values with the mean:***
```{r}
newdata <- data
newdata[is.na(data$steps), ]$steps <- mean(data$steps,na.rm=TRUE)
```

3. ***Create a new dataset:***
```{r}
newdata[1:10,]  #Print first ten lines
```

4. ***Create a histogram and calculate the mean and median of the new dataset:***
```{r}
totalstepsnew <- aggregate(steps ~ date, newdata, sum)
hist(totalstepsnew$steps)
```

Calculate the mean and median steps:
```{r}
mean(totalstepsnew$steps)
median(totalstepsnew$steps)
```

### Mean comparison:
The average number of  steps per day of the original dataset is ***`r mean(totalsteps$steps, na.rm=TRUE)`*** and the average of the imputed
dataset is ***`r mean(totalstepsnew$steps)`***. The values are similar.

### Median comparison:
The median number of  steps per day of the original dataset is ***`r median(totalsteps$steps, na.rm=TRUE)`*** and the median of the imputed
dataset is ***`r median(totalstepsnew$steps)`***. The values are similar.


## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
newdata$date <- as.Date(newdata$date, "%Y-%m-%d")
newdata$day <- weekdays(newdata$date)
newdata$type <- c("Weekday")
for (i in 1:nrow(newdata)){
  if (newdata$day[i] == "Saturday" || newdata$day[i] == "Sunday"){
    newdata$type[i] <- "Weekend"
  }
}
newdata$type <- as.factor(newdata$type)
aggdatatype <- aggregate(steps ~ interval+type, newdata, mean)
qplot(interval, steps, data=aggdatatype, geom=c("line"), xlab="Interval", 
      ylab="Average Steps per Day", main="Comparison of Average Steps on Weekdays and Weekends") + facet_wrap(~ type, ncol=1)
```

